Sweat electrolytes and CFTR activity in the era of CFTR modulators

### 2025-08-12 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
added method and filtered Age > 50 - need to figure out the multiple methods issue

@@

@@

@@

@@

@@

    labels = c("M/M", "M/S", "M/U", "M/S", "S/S", "S/U", "S/U", "U", "Not CF")
## By function
df$geno1_class <- cf_genotype_class_severity(df$Genotypes1, out = "class")
df$geno1_class[grep("508", df$Genotypes1)] <- "F508del"
df$geno1_class <- replace(
    df$geno1_class,
    df$geno1_class %in% c("I", "II", "III"),
    "MF"
df$geno1_class <- replace(
    df$geno1_class,
    df$geno1_class %in% c("IV", "V"),
    "RF"
df$geno2_class <- cf_genotype_class_severity(df$Genotypes2, out = "class")
df$geno2_class[grep("508", df$Genotypes2)] <- "F508del"
df$geno2_class <- replace(
    df$geno2_class,
    df$geno2_class %in% c("I", "II", "III"),
    "MF"
df$geno2_class <- replace(
    df$geno2_class,
    df$geno2_class %in% c("IV", "V"),
    "RF"
df$geno1_class[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno2_class[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno_class <- paste(df$geno1_class, df$geno2_class, sep = "/")
df$geno_class <- factor(
    df$geno_class,
        "F508del/F508del",
        "F508del/MF",
        "F508del/RF",
        "F508del/Unknown",
        "MF/F508del",
        "MF/MF",
        "MF/RF",
        "MF/Unknown",
        "RF/F508del",
        "RF/MF",
        "RF/Unknown",
        "Unknown/F508del",
        "Unknown/MF",

@@

        "MF/MF",
        "MF/MF",
        "MF/RF",
        "MF/U",
        "MF/MF",
        "MF/MF",
        "MF/RF",
        "MF/U",
        "MF/RF",
        "MF/RF",
        "RF/U",
        "MF/U",
        "MF/U",
        "U",

@@

# Organize columns

@@

    select(Name:Genotypes2, geno_risk, geno_class, everything()) %>%
    select(-geno_risk_1, -geno_risk_2, -geno1_class, -geno2_class) %>%
    rename(`Genotype Risk` = geno_risk, `Genotype Class` = geno_class)

@@

    numeric.stats = c("N", "Nmiss", "meansd", "medianq1q3", "range")

@@

```{r results='asis'}
summary(tableby(
    Diagnosis ~ Sex + Age + Na + Cl + `Na:Cl` + K,
    data = df,
    control = mycontrols
))
```

## Table 2: Participant Demographics by Genotype Class

In the table below "MF" stands for minimal function, "RF" for residual function, and "U" indicates unknown.

```{r results='asis'}
summary(tableby(
    `Genotype Class` ~ Sex + Age + Na + Cl + `Na:Cl` + K,
    data = df,
    control = mycontrols
))
```

## Table 3: Participant Demographics by Genotype Risk

In the table below "M" stands for mild, "S" for severe, and "U" indicates unknown.

```{r results='asis'}
summary(tableby(
    `Genotype Risk` ~ Sex + Age + Na + Cl + `Na:Cl` + K,
    data = df,
    control = mycontrols
))
```

@@

#### By genotype class

```{r}
#| warning: false
clp <- ggplot(df, aes(x = `Genotype Class`, y = Cl)) +
    geom_boxplot() +
    theme_bw()
clp
```

@@

#### By genotype class

```{r}
#| warning: false
nap <- ggplot(df, aes(x = `Genotype Class`, y = Na)) +
    geom_boxplot() +
    theme_bw()
nap
```

@@

#### By genotype class

```{r}
#| warning: false
ratiop <- ggplot(df, aes(x = `Genotype Class`, y = `Na:Cl`)) +
    geom_boxplot() +
    geom_hline(yintercept = 1, color = "red") +
    geom_hline(yintercept = 1.25, color = "red") +
    theme_bw() +
    ylab("Na:Cl") +
    xlab("Modulator")
ratiop
ratiop <- ggplot(df, aes(x = `Genotype Class`, y = `Na:Cl`)) +
    geom_boxplot() +
    geom_hline(yintercept = 1, color = "red") +
    geom_hline(yintercept = 1.25, color = "red") +
    ylim(0, 2) +
    theme_bw() +
    ylab("Na:Cl") +
    xlab("Modulator")
ratiop
```

>>>>

        Method,

@@

        Test.Method,

@@

        Method = Test.Method,

@@

method <- df %>%
    select(Name, Age, Method, Na, Cl) %>%
    drop_na()

@@

    select(-Method) %>%

@@

@@

        "Mild",
        "Mild",
        "Unknown",
        "Mild",
        "Severe",
        "Unknown",
        "Unknown",
        "Unknown",

@@

# Create a categorical column for Cl
df$`Cl Group` <- cut(
    df$Cl,
    breaks = c(-Inf, 30, 60, Inf),
    right = F,
    labels = c("< 30", "30-59", ">= 60")
# Add method back in
df <- left_join(df, method)
# Organize columns, remove age over 50 (there shouldn't be any age < 2 days)

@@

    select(Name:Genotypes2, geno_risk, everything()) %>%
    select(-geno_risk_1, -geno_risk_2) %>%
    rename(`Genotype Risk` = geno_risk) %>%
    filter(Age < 50)

@@

    numeric.stats = c("N", "Nmiss", "meansd", "medianq1q3", "iqr", "range")

@@

```{r results='asis'}
summary(
    tableby(
        Diagnosis ~
            Method +
                Sex +
                Age +
                fe(`Genotype Risk`) +
                fe(`Cl Group`) +
                Cl +
                Na +
                `Na:Cl` +
                K,
        data = df,
        simulate.p.value = T,
        control = mycontrols
    ),
    pfootnote = T
)
```

## Table 2: Participant Demographics by Genotype Risk

```{r results='asis'}
summary(
    tableby(
        `Genotype Risk` ~
            Sex +
                Age +
                fe(`Cl Group`) +
                Cl +
                Na +
                `Na:Cl` +
                K,
        data = df[
            df$`Genotype Risk` != "Unknown" & df$Diagnosis == "Cystic Fibrosis",
        ],
        simulate.p.value = T,
        control = mycontrols
    ),
    pfootnote = T
)
```

## Table 3: Participant Demographics by Chloride Group

```{r results='asis'}
summary(
    tableby(
        `Cl Group` ~ Sex + Age + Cl + Na + `Na:Cl` + K,
        data = df,
        control = mycontrols
    ),
    pfootnote = T
)
```

@@

@@

@@
